Doing Better Than UCT: \\ Rational Monte Carlo Sampling in Trees

### 2011-08-03 paper/rmcts/paper.tex
initial draft version

Such sampling scheme is significantly less sensitive to the tuning of
the empirical upper bound ($C_p$ in \cite{Kocsis.uct}) of UCT, since
the conflict \cite{Bubeck.pure} between the need for a larger value of $C_p$ on the first
step (simple regret) and a smaller value for the rest of the playout
(accumulated regret) is resolved. In fact, a sampling scheme that uses
UCB at all steps but a larger value of $C_p$ for the first step than
for the rest of the steps, outperforms UCT.

@@

@@

@@

@@

@@

@@

>>>>

Such sampling scheme is significantly less sensitive to the tuning of
the exploration factor ($C_p$ in \cite{Kocsis.uct}) of UCT, since
the conflict \cite{Bubeck.pure} between the need for a larger value of $C_p$ on the first
step (simple regret) and a smaller value for the rest of the playout
(accumulated regret) is resolved. In fact, a sampling scheme that uses
UCB at all steps but a larger value of $C_p$ for the first step than
for the rest of the steps, outperforms UCT.

@@

Figure~\ref{fig:mab-simple-regret} presents a comparison of simple
regret minimization algorithms for Multi-armed
bandits. Figure~\ref{fig:mab-simple-regret}.a shows the search tree
corresponding to a problem instance. Each arm returns a random reward
drawn from a Bernoulli distribution. The search selects an arm
and compares the expected reward, unknown to the algorithm during the
sampling, to the expected reward of the best arm.

Figure~\ref{fig:mab-simple-regret}.b shows the regret
vs. the number of samples, averages over $10^4$ experiments, for uniform
random sampling (RND), UCT (same as UCB in this case),
1/2-greedy (GCT), and QCT (UCT with $\log(\cdot)$ replaced by
$\sqrt{\cdot}$), for randomly generated instances of 32 arms.

For smaller numbers of samples, 1/2-greedy achieves the best
performance; for large number of samples, QCT outperforms GCT. QCT is
better than UCT everywhere except for very small numbers of samples. A
combination of GCT and QCT dominates UCT over the whole range.

@@

The second set of experiments was performed on randomly generated
trees crafted in such a way that uniform random sampling selects a
direction at the root randomly. The degree of the root is a parameter of
the tree generator. The degree of all nodes at distance 1 from the
root is 2, and all nodes at distance 2 from the roots are leaves. The
average reward of two children of each node at distance 1 is
0.5. Thus, a uniform sampling scheme results in the same average reward for
all edges at the root, and an adaptive sampling scheme, such as UCT,
has to be used.

Figure~\ref{fig:mcts-regret} shows a sketch of the search tree
(Figure~\ref{fig:mcts-regret}.a) and the dependency of the regret vs. the
number of samples for trees with root degree 16
(Figure~\ref{fig:mcts-regret}.b) and 64 (Figure~\ref{fig:mcts-regret}.c). The
dependencies look differently from Multi-armed bandit instances, but
the algorithms exhibit a similar relative performance: GCT and QCT
simultaneously are better than UCT, QCT dominates UCT everywhere
except for small numbers of instances. The advantage of QCT grows with
the number of arms.

@@

Figures~\ref{fig:sailing-cost-vs-nsamples}--\ref{fig:sailing-rcq-vs-factor}
show results of experiments on the sailing
domain. Figure~\ref{fig:sailing-cost-vs-nsamples} shows the regret
vs. the number of samples, computed for a range of values of
$C_p$. Figure~\ref{fig:sailing-cost-vs-nsamples}.a shows the median
cost, and Figure~\ref{fig:sailing-cost-vs-nsamples}.b --- the minimum
costs. UCT is always worse than either 1/2-greedy (GCT) or QCT, and is sensitive to
the value of $C_p$: the median cost is much higher than the minimum
cost for UCT. For both 1/2-greedy and QCT, the difference is
significantly less prominent.

@@

Figure~\ref{fig:sailing-cost-vs-factor} shows the rerget vs. the
factor for different numbers of samples. QCT is always better than
UCT, and 1/2-greedy is better than UCT expect for a small range of
values of the exploration factor $C_p$.

@@

Figure~\ref{fig:sailing-lake-size} shows the cost vs. the exploration
factor for lakes of different sizes. The relative difference between
the sampling schemes becomes more prominent when the lake size
increases.

@@

Figure~\ref{fig:sailing-rcq-vs-factor} compares QCT with UCT with a
different exploration factor at the root (CCT). $C_p$ for the rest of
the steps was chosen to ensure the best performance from earlier
experiments on the domain. Both algorithms exhibit similar dependency,
but as the number of samples grows, QCT achieves smaller average
regrets and is less sensitive to the choice of the value for $C_p$ at
the first step.
